A Switching Planner for Combined Task and Observation Planning

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
Incorporated some of Charles' suggestions.

Planning for agents acting in partially observable environments with
stochastic actions is extremely computationally
expensive \cite{mdp-complexity}. Even relatively small partially
observable Markov decision problems (POMDPs), represented by tens of
propositional variables, produce state spaces far larger than existing
optimal or close to optimal planners can handle. For the easier
problem of planning in completely observable stochastic domains, an
alternative approach has been to use classical sequential planners and
replan when actions have unplanned-for
outcomes \cite{yoon:etal:2007}. However, these approaches aren't so
useful in partially observable domains as replanning is triggered when
the system ends up in a state other than the desired one, and this
state can't in general be determined.

In this paper we describe, \pcogx, the planning component in a robotic
system that continuously deliberates in a stochastic dynamic
environment in order to achieve objectives set by the user, and
acquire knowledge about its surroundings. This domain has the
characteristics of a POMDP in that the state is not perfectly known,
and state information is gained by the robot through the use of
unreliable sensing actions. However, the state representation is
propositional, and even for relatively simple problems is too large
for a POMDP solver to be applied.

Inspired by the replanning approaches used successfully for MDP
planning \cite{yoon:etal:2007,yoon:etal:2008} we propose an approach
that uses a replanning sequential planner to build the majority of the
plan, but uses a decision-theoretic planner on small subproblems where
reasoning about observations is required to find good plans. We refer
to the complete system as a {\em switching planner}. The approach is
not optimal, particularly as it is based on a satisficing sequential
planner, but it performs better than the sequential planner alone and
is fast enough to be used onboard the robot for decision-making. As an
illustration, consider a robot searching for a box that it knows is
often found in offices. The sequential planner might produce a plan
that assumes that {\sc Room a} is an office, that the box is in {\sc
Room a}, and that the box is in location 1 in {\sc Room a}, and then
goes to {\sc Room a}, goes to location 1 and searches for the
box. When the plan gets executed and the search begins, the
decision-theoretic planner is called to conduct the search, and at
this point may retract the last assumption and build a plan to search
the whole of {\sc Room a} for the box. If this plan fails, it will do
so by disproving one of the assumptions (for example, that {\sc Room
a} is an office), which will then lead to replanning by the sequential
planner based on the newly discovered knowledge.

To represent planning problems of this kind, we have developed a
first-order declarative language, called decision-theoretic planning
domain definition language (DTPDDL), for describing domains of
planning problems that correspond to POMDPs. We have created a system
that automatically constructs from this a determinised planning domain
for the sequential planner that includes actions that allow
assumptions to be made about the values of imperfectly known state
variables. These are then used to restrict the state space and set the
reward function for the decision-theoretic planner, which can operate
with the original representation of the domain.

The sequential planner we use is Continual
Planning \cite{brenner:nebel:jaamas09} (CP). CP is based
on \fastdownward \cite{fast-downward}. CP already has the capability
to replan, and also allows agent knowledge to be represented
explicitly, so we can write actions that gain the agent knowledge. For
the decision-theoretic planner we use an information-state
\laostar\ procedure that uses the solution to the equivalent MDP as a
heuristic. For reasons of space, we will not describe this system in
detail here.

The switching planner approach gives us some important advantages in
our robotic domain. Firstly, the main source of uncertainty in the
domain is unreliable observations, rather than stochasticity in the
action effects, and this approach is particularly suited to problems
that combine a deterministic task planning problem with observations
as the decision-theoretic planner is only invoked to determine the
values of variables. Secondly, the replanning makes the system very
robust to changing goals and the discovery of new facts about the
world, both of which are features of our domain.

>>>>

Planning for agents acting in partially observable environments with
stochastic actions is extremely computationally
expensive \cite{mdp-complexity}. Even relatively small partially
observable Markov decision problems (POMDPs), represented by tens of
propositional variables, produce state spaces far larger than existing
optimal or close to optimal planners can handle. For the easier
problem of planning in completely observable stochastic domains, an
alternative approach has been to use classical sequential planners and
replan when actions have unplanned-for
outcomes \cite{yoon:etal:2007}. However, these approaches rely
crucially on being able to determine properties of the current state
in order to decide to replan. For partially observable domains, only
probabilistic information about these properties will in general be
available.

In this paper we describe, \pcogx, the planning component in a robotic
system that continuously deliberates in a stochastic dynamic
environment in order to achieve objectives set by the user, and
acquire knowledge about its surroundings. This domain has the
characteristics of a POMDP in that the state is not perfectly known,
and state information is gained by the robot through the use of
unreliable sensing actions. However, the state representation is
propositional, and even for relatively simple problems the underlying
model is too large for a POMDP solver to be applied.

Inspired by the replanning approaches used successfully for MDP
planning \cite{yoon:etal:2007,yoon:etal:2008} we propose an approach
that uses a replanning sequential planner to build the majority of the
plan, but uses a decision-theoretic planner on small subproblems where
reasoning about observations is required to find good plans. We refer
to the complete system as a {\em switching continual planner}. The
approach is not optimal, particularly as it is based on a satisficing
sequential planner, but it performs better than the sequential planner
alone and is fast enough to be used in real-time on a mobile robot for
decision-making.

As an illustration, consider a robot searching for a box that it knows
is often found in offices. The sequential planner might produce a plan
that assumes that {\sc Room a} is an office, that the box is in {\sc
Room a}, and that the box is in location 1 in {\sc Room a}. Given
these assumptions, the plan might be to go to {\sc Room a}, go to
location 1 and search for the box. When the plan gets executed and the
search begins the decision-theoretic planner is called since searching
involves making observations. The decision-theoretic planner might
retract the last assumption but leave the others, and build a plan to
search the whole of {\sc Room a} for the box. If this plan fails to
find the box, it will disprove the assumption that the box is in {\sc
Room a}, which will then lead to replanning by the sequential planner
based on the newly discovered knowledge.

To represent planning problems of this kind, we have developed a
first-order declarative language, called decision-theoretic planning
domain definition language (DTPDDL), for describing domains of
planning problems that correspond to POMDPs. We have created a system
that automatically constructs from this a deterministic planning domain
for the sequential planner that includes actions that allow
assumptions to be made about the values of imperfectly known state
variables. These are then used to restrict the state space and set the
reward function for the decision-theoretic planner, which can operate
with the original representation of the domain.

The sequential planner we use is based
on \fastdownward \cite{fast-downward}. We add the capability to
replan, and also allow agent knowledge to be represented explicitly,
so we can write actions that gain the agent knowledge. For the
decision-theoretic planner we use forward search in the information
state.

The switching planner approach gives us some important advantages in
our robotic domain. First, the main source of uncertainty in the
domain is unreliable observations, rather than stochasticity in the
action effects, and this approach is particularly suited to problems
that combine a deterministic task planning problem with observations
as the decision-theoretic planner is only invoked to determine the
values of variables. Secondly, the replanning makes the system very
robust to changing goals and the discovery of new facts about the
world, both of which are features of our domain.
